---
title: "TbasCO - The Basics"
author: "Joris van Steenbrugge"
date: "October 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set()
```

#Loading the package

First we start by loading the package. By default, when TbasCO is loaded, the URLs of a sample dataset and annotation database is added to the environment as two separate variables `file.path` and `annotation.db.path`. If you are using Rstudio you can look at the environment tab before and after we run the following chunk of code. These two variables may be replaced by your own dataset.

```{r}
library(TbasCO)
```


#Loading and Preprocessing Data


All of the preprocessing is handled by the `Pre_process_input` function.
This function consecutively performs the following tasks:

1. Data Loading
2. Normalization -> Default normalization method between samples and by bin
3. Filtering -> the choice between Mean Absolute Deviation (MAD) and based on Standard Deviation

```{r}
RNAseq.data <- Pre_process_input(file.path,
                                 annotation.db.path,
                                 normalize.method    = T,
                                 filter.low.coverage = T)
```


### Actual Analysis
```{r Main Analysis, message=FALSE, warning=FALSE, include=FALSE, results=}
Trait.attributes <- GetTraitAttributes(RNAseq.data = RNAseq.data,
                                       metrics = distance.metrics, N =)

```


### Using a species as a model
```{r}
# The name of the Bin/genome, available options can be seen by running: RNAseq.data$features$bins
my_model <- '39'

# Extract the name of the traits that are present in 'my_model'

Model_Module_List_All <- Model_Module(RNAseq.data, trait.attributes, 39, bkgd.traits)
Plot_Model_Module(Model_Module_List_All, 39, Module_Names,margins, sortbygenome="16")

```
